Article title

### 2017-07-21 paper-webserver.tex
ADD: bunch of new paras wrt Chads suggestions, still need to expand on drawbacks

Amongst produce in Canada, \textit{Salmonella enterica} and \textit{Escherichia coli} are the two most common bacetrial infections \cite{kozak2013foodborne}.
Translate into actual costs: hospitalization adn death \cite{thomas2015estimates}.
Of which, monetary costs have been known for decades \cite{todd1989costs}.

The surveillance of bacterial pathogens is a routine function of public-health agencies.
In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are becoming displaced by culture-independent methods based on whole-genome sequencing (WGS).
Secondary analyses, the study of relations and historical relevance of genomes, are now possible on large datasets provided by WGS \cite{ronholm2016navigating}.
Advances in WGS technologies have also enabled the monitoring and sharing of outbreak genomes in real-time \cite{ronholm2016navigating}.

Subtyping and other analyses software often require command-line or programming experience.
To make such \textit{in-silico} techniques user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
While useful, arguably \cite{fricke2014bacterial}, the lack of bioinformatic standards and common infrastructure technologies have made it difficult to fully leverage the big-data results from WGS.

Previously we developed Superphy \cite{whiteside2016superphy}, an online platform targeting \textit{E. coli}, which determines Shiga toxin subtypes, Antimicrobial Resistance (AMR) genes using the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and provides statistical significance tests between subgroups in associated metadata.
Here we present Spfy, a complete rewrite of Superphy integrating our Panseq package \cite{laing2010pan} for pan-genomic analyses of accessory regions, long-term storage of results in a graph database, and containerization of services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
Furthermore, Spfy takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel.
All services are publicly available at \url{https://lfz.corefacility.ca/superphy/spfy/}.

>>>>

% costs of bacterial infections
With the prevalence of ready-to-eat produce and expansion of global food-supply chains, consumers are a greater risk for bacterial infections in an environment that is more complicated to track than ever before \cite{kozak2013foodborne}.
Amongst produce in Canada, \textit{Salmonella enterica} and \textit{Escherichia coli} are the two most common bacetrial infections \cite{kozak2013foodborne}.
Foodborne illnesses translate into actual costs: in Canada alone, an estimated 4000 hospitalizations and 105 deaths are thought to be the result of various pathogens \cite{thomas2015estimates}.
Futhermore, economic costs have been known for decades \cite{todd1989costs}.

% surveillance
The surveillance of bacterial pathogens is a routine function of public-health agencies in an effort to lower the burden caused by human illness.
In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are lately coupled with culture-independent methods based on whole-genome sequencing (WGS) to monitor known pathogens and detect emerging threats.

% traditional reference lab techniques
Subtyping encompasses various methods to designate particular strains of bacterial isolates.
In order to monitor outbreaks, platforms must be able to rapidly subtype collected strains and communicate such information to other agencies \cite{ronholm2016navigating}. 
Traditional reference laboratory techniques handled subtyping through methods such as serotyping, detecting the presence of specific cell surface antigens as tested by antibody reaction, pulsed-field gel electrophoresis (PFGE), to identify known virulence factors (VFs) or antimicrobial resistance genes (AMR genes), or Sanger sequencing, to determine multi-locus sequence types (MLSTs) \cite{ronholm2016navigating}.
Using PFGE, outbreaks can be tracked by comparing band patterns of bacterial isolates, and is routinely performed in public health agencies \cite{kozak2013foodborne}. 
The original computational methods to leverage WGS focused on creating \textit{in-silico} approaches to return subtyping results matching that of wet-lab techniques.
Software such as VirulenceFinder \cite{joensen2014real} compares a pre-computed set of reference sequences encompassing known virulence factor gene sequences against a WGS result.
Similar approaches have been applied to predict AMR genes \cite{mcarthur2013comprehensive}, O-type, and H-type \cite{whiteside2016superphy}.

% potential of WGS
WGS has greater potential due to the inclusion of the entire genomes; in effect, this lends WGS-based methods better resolution and context in evaluating the spread of outbreaks \cite{ronholm2016navigating}.
For example, instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes, as done for laboratory-based MLST, newer methods such as core genome MLST (cgMLST) attempt to cover every common gene.
Secondary analyses, the study of relations and historical relevance of genomes, are now possible on large datasets provided by WGS \cite{ronholm2016navigating}.
Advances in WGS technologies have also enabled the monitoring and sharing of outbreak genomes in real-time \cite{ronholm2016navigating}.

% Drawbacks
However, these WGS-base methods still examine the underlying genome, not necessarily the expressed genes.
Correspondingly, a greater number of target sequences are identified which complicates risk assessments \cite{ronholm2016navigating}.
The effectiveness of WGS-based methods is also largely reliant on the accuracy and breadth of reference datasets.

% Solution via pre-computed analyses for routine tasks
Efforts such as the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive} attempt to alleviate data concerns by curating a set of known AMR genes.
In addition, agencies such as the Food and Drug Administration (FDA) (CITE GENOME TRACKR) and the National Center for Biotechnology Information (NCBI) (CITE ???) provide services to openly share genome sequences and associated metadata.
Recent efforts have been focused on providing modern tools for quality control of reference genomes (CITE SOME PAPER ON QC OF BACETIRA GENOMES), particularly amongst human-genome projects (CITE NGSCHECKMATE, GOOGLE GENOMICS).
Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, and chewBBACA \url{https://github.com/mickaelsilva/chewBBACA} are available to compute core and accessory regions and identify SNPs across a large collection of reference genomes.
Taken together, open-access initiatives and shared tooling aim to alleviate the drawbacks of WGS-based methods by pre-computing sufficient reference data.

% why Superphy
Subtyping and computing reference data often require command-line or programming experience.
To make such \textit{in-silico} techniques user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
While useful, arguably \cite{fricke2014bacterial}, the lack of bioinformatic standards and common infrastructure technologies have made it difficult to fully leverage the big-data results from WGS.

% Superphy
Previously we developed Superphy \cite{whiteside2016superphy}, an online platform targeting \textit{E. coli}, which determines Shiga toxin subtypes, Antimicrobial Resistance (AMR) genes using CARD, and provides statistical significance tests between subgroups in associated metadata and pan-genome regions.
However, we struggled to provide real-time computing of results and to accomdate an ever-growing reference database.
Furthermore, due to the nature of the various dependencies involved, the platform was difficult for other researchers to independently replicate.

% this study
Here we present Spfy, a complete rewrite of Superphy which expands on supported features to also provide long-term storage of results in a graph database, and containerization of services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
Furthermore, Spfy takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel.
These changes enable us to provide real-time genome analysis and database storage of all results for upwards of 5000 genomes.
All services are publicly available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
